---
title: "Modeling Senior Theses"
output:
  xaringan::moon_reader:
    css: ["fc", "fc-fonts", "reed.css", "default"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
library(tidyverse)
library(infer)
```

# Case study: senior theses

```{r echo = TRUE}
theses <- read.csv("../data/sample_theses.csv")
str(theses)
```


---
class:small
**Question 1**: What is the average number of times a thesis is checked out? (Description)

--

```{r, fig.height=3, fig.width=5, fig.align='center'}
ggplot(theses, aes(x = checkouts)) +
  geom_bar() +
  theme_bw(base_size = 18) 
```

--

```{r}
theses %>%
  summarize(mean(checkouts),
            median(checkouts))
```


---
class:small
**Question 1**: What is the average number of times a thesis is checked out? (Inference) 

--

**Confidence interval on one mean (approximation method)**

$$
\bar{x} \pm t^* \times s/\sqrt n
$$
--

```{r}
stats <- theses %>%
  summarize(n     = n(),
            x_bar = mean(checkouts),
            s     = sd(checkouts))
stats$x_bar + c(1, -1) * qt(.025, df = stats$n - 1) *
  stats$s / sqrt(stats$n)
```


---
class:small
**Question 1**: What is the average number of times a thesis is checked out? (Inference) 

**Confidence interval on one mean (computational method)**

--

```{r}
se_boot <- theses %>%
  specify(response = checkouts) %>%
  generate(500, type = "bootstrap") %>%
  calculate(stat = "mean") %>%
  summarize(sd(stat)) %>% 
  pull()
```

--

```{r}
stats$x_bar + c(-1, 1) * 2 * se_boot
```


---
class:small
**Question 2**: Is the distribution of theses uniform in time? (Description)

--
```{r, fig.height = 5}
ggplot(theses, aes(x = year)) +
  geom_histogram() +
  theme_bw(base_size = 18)
```


---
class:small
**Question 2**: Is the distribution of theses uniform in time? (Inference)

--

**Chi-squared goodness of fit test **

```{r, fig.width=7, fig.height=4, eval = FALSE}
bd <- data.frame(year_binned =
                   cut(theses$year, seq(from = 1920, 
                                        to = 2020, 
                                        by = 10)))
ggplot(bd, aes(x = year_binned)) +
  geom_bar() +
  geom_hline(yintercept = stats$n/10, 
             color = "goldenrod", lwd = 2) +
  theme_bw(base_size = 18)
```


---
class:small
**Question 2**: Is the distribution of theses uniform in time? (Inference)

**Chi-squared goodness of fit test **

```{r, fig.width=7, fig.height=6, echo = FALSE}
bd <- data.frame(year_binned =
                   cut(theses$year, seq(from = 1920, 
                                        to = 2020, 
                                        by = 10)))
ggplot(bd, aes(x = year_binned)) +
  geom_bar() +
  geom_hline(yintercept = stats$n/10, 
             color = "goldenrod", lwd = 2) +
  theme_bw(base_size = 18)
```

---
class:small
**Question 2**: Is the distribution of theses uniform in time? (Inference)

**Chi-squared goodness of fit test **

--

```{r}
(obs <- table(bd$year_binned))
```

--

```{r}
(exp <- length(bd$year_binned)/10)
```

--

```{r}
(chisq_obs <- sum((obs - exp)^2/exp))
```

---
class:small
**Question 2**: Is the distribution of theses uniform in time? (Inference)

**Chi-squared goodness of fit test (approximation method) **

```{r}
pchisq(chisq_obs, df = length(obs) - 1, lower.tail = FALSE)
```


---
class:small
**Question 2**: Is the distribution of theses uniform in time? (Inference)

**Chi-squared goodness of fit test (computational method) **

```{r}
levels(bd$year_binned) <- 1:10
null <- bd %>%
  specify(response = year_binned) %>%
  hypothesize(null = "point", p = c("1" = .1, 
                                    "2" = .1,
                                    "3" = .1,
                                    "4" = .1,
                                    "5" = .1,
                                    "6" = .1,
                                    "7" = .1,
                                    "8" = .1,
                                    "9" = .1,
                                    "10" = .1)) %>%
  generate(500, type = "simulate") %>%
  calculate(stat = "Chisq")
```


---
class:small
**Question 2**: Is the distribution of theses uniform in time? (Inference)

**Chi-squared goodness of fit test (computational method) **

```{r fig.height = 4.5}
library(ggthemes)
visualize(null) +
  shade_p_value(obs_stat = chisq_obs, direction = "right") +
  theme_economist()
```


---
class:small
**Question 3**: What is the relationship between the age of a thesis and the number of checkouts? (Description)

--

```{r}
head(theses)
```

--

```{r}
theses <- theses %>%
  mutate(age = 2017 - year)
head(theses)
```


---
class:small
**Question 3**: What is the relationship between the age of a thesis and the number of checkouts?  (Description)

```{r fig.height = 5}
ggplot(theses, aes(x = age, y = checkouts)) +
  geom_jitter() +
  theme_bw(base_size = 18)
```

---
class:small
**Question 3**: What is the relationship between the age of a thesis and the number of checkouts? (Inference)

--

```{r}
m1 <- lm(checkouts ~ age, data = theses)
summary(m1)
```

---
class:small
**Question 3**: What is the relationship between the age of a thesis and the number of checkouts? (Inference)

```{r fig.height = 5}
ggplot(theses, aes(x = age, y = checkouts)) +
  geom_jitter() +
  theme_bw(base_size = 18) +
  stat_smooth(method = "lm", se = FALSE, lwd = 2)
```


---

boardwork



---
class:small
# Poisson Regression
--

```{r}
m2 <- glm(checkouts ~ age, data = theses, family = "poisson")
summary(m2)
```

---
class:small

```{r fig.height = 5}
ggplot(theses, aes(x = age, y = checkouts)) +
  geom_jitter() +
  theme_bw(base_size = 18) +
  stat_function(fun = function(age) {exp(coef(m2)[1] + 
                                           coef(m2)[2] * age)},
                color = "red", lwd = 2)
```

---
class:small

```{r}
t2 <- theses %>%
  filter(year > 1994)
m2 <- glm(checkouts ~ age, data = t2, family = "poisson")
summary(m2)
```

---
class:small
```{r fig.height = 5}
ggplot(t2, aes(x = age, y = checkouts)) +
  geom_jitter() +
  theme_bw(base_size = 18) +
  stat_function(fun = function(age) {exp(coef(m2)[1] + 
                                           coef(m2)[2] * age)},
                color = "red", lwd = 2)
```

<!-- --- -->
<!-- # From the library's database... -->

<!-- -- -->

<!-- ```{r} -->
<!-- theses_lib <- read.csv("../data/library_theses.csv") -->
<!-- head(theses_lib) -->
<!-- ``` -->


<!-- --- -->

<!-- **Question 1**: What is the average number of times a thesis is checked out? -->

<!-- ```{r, fig.height=3, fig.width=5, fig.align='center'} -->
<!-- qplot(x = checkouts, data = theses_lib, geom = "bar") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- theses_lib %>% -->
<!--   summarize(mean(checkouts), -->
<!--             median(checkouts)) -->
<!-- ``` -->


<!-- --- -->

<!-- **Question 2**: Is the distribution of theses uniform in time? -->

<!-- ```{r} -->
<!-- qplot(x = year, data = theses_lib, geom = "histogram") -->
<!-- ``` -->


<!-- --- -->

<!-- **Question 3**: What is the relationship between the age of a thesis and the number of checkouts? -->

<!-- ```{r} -->
<!-- theses_lib <- theses_lib %>% -->
<!--   mutate(age = 2017 - year) -->
<!-- qplot(x = age, y = checkouts, data = theses_lib, geom = "jitter") -->
<!-- ``` -->


<!-- --- -->
<!-- # Poisson Regression -->
<!-- -- -->

<!-- ```{r} -->
<!-- t3 <- theses_lib %>% -->
<!--   filter(year > 1994) -->
<!-- m3 <- glm(checkouts ~ age, data = t3, family = "poisson") -->
<!-- summary(m3) -->
<!-- ``` -->


<!-- --- -->

<!-- ```{r} -->
<!-- qplot(x = age, y = checkouts, data = t3, geom = "jitter") + -->
<!--   stat_function(fun = function(age) {exp(coef(m3)[1] + coef(m3)[2] * age)}, -->
<!--                 color = "red") -->
<!-- ``` -->

<!-- --- -->
<!-- # How'd we do? -->

<!-- ```{r} -->
<!-- coef(m2) # our data -->
<!-- coef(m3) # library's data -->
<!-- ``` -->


