---
title: "Hypothesis Testing Review"
output:
  xaringan::moon_reader:
    css: ["fc", "fc-fonts", "reed.css", "default"]
    lib_dir: libs
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
library(knitr)
options(digits=2)
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```


# What is it good for?

--

Hypothesis test have been shown to be valuable contributors to science (p < .05)
but are sometimes abused (p < .05).

--

- Used to assess the degree to which data is consistent with a particular model.
- The most widely used tool in statistical inference.


---
# Step 1

--

Lay out your model(s).

$H_0$: null model, business as usual  
$H_A$: alternative model, business not as usual

--

- Hypotheses are statments about the TRUE STATE of the world and should involve
*parameters*, not *statistics*.
- Hypotheses should suggest a *test statistic* that has some bearing on the claim.
- Always use two-tailed tests.


---
# Step 2

--

Contruct the appropriate null distribution.

```{r, echo=FALSE, eval=TRUE, fig.height = 3.5, fig.width = 10, fig.align='center'}
x1  <- 0:75
df <- data.frame(x = x1, y = dbinom(x1, 75, 0.5))
qplot(x, y, data = df) + 
  geom_bar(stat = "identity", col = "darkgreen", fill = "white") +
  stat_function(fun = dnorm, args = list(mean = 75/2, sd = sqrt(74/4)), 
                col = "goldenrod", lwd = 1.2)
```

--

1. Randomization
--

2. Simulation
--

3. Exact Probability Theory
--

4. Normal Approximation


---
# Step 3

--

Calculate a measure of consistency between the observed test statistic (the data)
and the null distribution (i.e., a p-value).

--

- If your observed test stat is in the tails > low p-val > data is inconsistent
with null hypothesis > "reject null hypothesis".
- If your observed test stat is in the body > high p-val > data is consistent with
the null hypothesis > "fail to reject the null hypothesis".

--

#### What can go wrong?

