---
title: "Lab 8: Inference for Numerical Data"
output: github_document
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(eval = FALSE)
library(tidyverse)
library(openintro)
library(infer)
```

## Getting Started

### Load packages

In this lab, we will explore and visualize the data using the **tidyverse** suite of 
packages, and perform statistical inference using **infer**. The data can be found
in the companion package for the textbook, **openintro**.

Let's load the packages.

```{r load-packages, message=FALSE}
library(tidyverse)
library(infer)
library(openintro)
```

### The data

Every two years, the Centers for Disease Control and Prevention conduct the Youth
Risk Behavior Surveillance System (YRBSS) survey, where it takes data from high 
schoolers (9th through 12th grade), to analyze health patterns. You will work 
with a selected group of variables from a random sample of observations during 
one of the years the YRBSS was conducted.

Load the `yrbss` data set into your workspace.

```{r load-data}
data(yrbss)
```

There are observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable can be found by bringing up the help file:

```{r help-nc}
?yrbss
```


1.  What are the cases in this data set? How many cases are there in our sample?

Remember that you can answer this question by viewing the data in the data viewer or
by using the following command:

```{r str}
glimpse(yrbss)
```

## Exploratory data analysis

You will first start with analyzing the weight of the participants in kilograms:
`weight`.

Using visualization and summary statistics, describe the distribution of weights.
The `summary` function can be useful.

```{r summary}
summary(yrbss$weight)
```

2.  How many observations are we missing weights from?

Next, consider the possible relationship between a high schooler's weight and their
physical activity. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

First, let's create a new variable `physical_3plus`, which will be coded as either
"yes" if they are physically active for at least 3 days a week, and "no" if not.

```{r create new var}
yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(physically_active_7d > 2, "yes", "no"))
```


3.  Make a side-by-side boxplot of `physical_3plus` and `weight`. Is there a 
relationship between these two variables? What did you expect and why?

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following to 
first group the data by the `physical_3plus` variable, and then calculate the mean
`weight` in these groups using the `mean` function while ignoring missing values
by setting the `na.rm` argument to `TRUE`.

```{r by-means}
yrbss %>%
  group_by(physical_3plus) %>%
  summarise(mean_weight = mean(weight, na.rm = TRUE))
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test.

## Inference Using the t-distribution

The classical method to address this question is to conduct a two-sample t-test.

4.  Conduct a two-sample t-test that includes the following steps:
  - Write the hypotheses for testing if the average weights are different for those
who exercise at least times a week and those who don't.
  - Compute the observed test statistic (you can `group_by()` and `summarize()` to compute things like group size (`n()`) and standard deviation (`sd()`)).
  - Find the degrees of freedom then use `pt()` to find the p-value associated with that observed statistic.
  - State your conclusions regarding the null hypothesis.


## Inference Using Permutation

An alternative technique is to skip the t-distribution and use a computational technique.

Let's start by calculating and saving the observed test statistic, which we'll call `obs_diff`.

```{r inf-weight-habit-ht-initial, tidy=FALSE, warning = FALSE}
obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

Notice how you have to specify the order of the difference in the same manner that you 
did when looking at the difference in proportions.

Next you can use permutation to generate 500 data sets under the notion that the `weight` is independent of `physical_3plus`.

```{r inf-weight-habit-ht-null, tidy=FALSE, warning = FALSE}
null <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 500, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))
```

We can visualize this null distribution with the following code:

```{r}
null %>%
  visualize()
```

5. How many of these `null` permutations have a difference of at least `obs_diff`? Shade those statistics in your null distribution plot and calculate the precise proportion (the p-value).

* * *

## More Practice

6.  Calculate a 95% confidence interval for the average height in meters (`height`)
and interpret it in context.

7.  Conduct a hypothesis test evaluating whether the average height is different
for those who exercise at least three times a week and those who don't.

8.  Now, a non-inference task: Determine the number of different options there 
are in the dataset for the `hours_tv_per_school_day` there are.

9. Come up with a research question evaluating the relationship between height 
or weight and sleep. Formulate the question in a way that it can be answered using
a hypothesis test and/or a confidence interval. Report the statistical results, 
and also provide an explanation in plain language. Be sure to check all 
assumptions, state your alpha level, and conclude in context.

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
</div>

* * *

# Problem Set

Please work through problems 5, 10, 14, 30, and 35 in the Exercises at the end of Chapter 4.